Bird beak's elaboration and innovation

### 2021-11-08 Analysis/03.1-results.Rmd
Update 03.1-results.Rmd

```{r, eval = TRUE, echo = FALSE, print = FALSE}
library(dispRity)
```

@@

 * `between`: the projection of each group's VCV (major axis) on the phylogenetic's VCV (major axis)
 * `phylo`: the projection of each element within a group (species) onto the phylogenetic's VCV (major axis)
 * `within`: the projection of each element within a group (species) onto its own VCV (major axis)

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = TRUE}
## Run the analyses between groups
super_results_between <- dispRity.covar.projections(superorder_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses between groups
order_results_between <- dispRity.covar.projections(order_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses of groups on the phylogeny
super_results_phylo <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Save all
shapespace_allbirds_lvl_superorder_order_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(shapespace_allbirds_lvl_superorder_order_results_list, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
```

@@

### Projections between groups:

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
## Plot the results
wrap.plot <- function(results, col, main, names, centre.position = TRUE, absolute = FALSE) {

@@

    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    op <- par(mfrow = c(1, n+1))
    ## Plot the first panel
    par(bty = "n", mar = c(5, 14, 4, 0) + 0.1)
    empty <- matrix(0, ncol = length(results[[1]]$disparity))
    colnames(empty) <- names(results[[1]]$disparity)
    boxplot(empty, horizontal = TRUE, las = 2, col = "white", ylim = c(0,0), xaxt = "n", border = "white")
    for(i in 1:n) {
        ## Is position?
        is_position <- names(results)[[i]] == "position"
        par(mar = c(5, 0, 4, 0) + 0.1, bty = "o")
        one.plot(results[[i]], col = col, centre = is_position && centre.position, main = ifelse(i == centre, main, ""), xlab = names_list[i], zero = is_position, absolute = absolute)
    }
    par(op)
}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[between] superorders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[between] superorders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

```{r, fig.height = 16, fig.width = 8, echo = FALSE}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[between] orders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[between] orders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot the results
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[phylo] superorders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[phylo] superorders species\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

## Plot the results
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[phylo] orders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[phylo] orders species\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

## Plot the results
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[within] superorders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[within] superorders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

## Plot the results
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[within] oreders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[within] oreders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

### Extreme species per super order (based on phylogeny)

@@

### Extreme species per order (based on phylogeny)

@@

### Extreme species per super order (within)

@@

### Extreme species per order (within)

@@

```{r, echo = FALSE}
get.correlations <- function(res, cent.tend = median, exclude.phylo = FALSE, absolute = TRUE, ...) {

@@

    ## Calculate and plot the hdrcde
    op_tmp <- par(mfrow = c(ceiling(sqrt(n_groups)), round(sqrt(n_groups))))
    for(i in 1:n_groups) {
        ## Testing the correlation
        cor_test <- cor.test(centrals_pos[[i]], centrals_dis[[i]])
        plot_main <- paste0(names(centrals_pos)[i], ": ", unname(round(cor_test$estimate, 3)), " (p=", unname(round(cor_test$p.value, 3)), ")")
        ## Plotting the results
        plot(centrals_pos[[i]], centrals_dis[[i]], pch = 19, xlab = "elaboration", ylab = "innovation", main = plot_main, cex = 0.5)
    }
    par(op_tmp)
    return(invisible())
}
```

### Correlations per super order (based on phylo)

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
get.correlations(results$super_results_phylo)
```

### Correlations within super orders

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
get.correlations(results$super_results_within, exclude.phylo = TRUE)
```

### Correlations per order (based on phylo)

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
get.correlations(results$order_results_phylo, exclude.phylo = TRUE)
```

### Correlations within orders

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
get.correlations(results$order_results_within, exclude.phylo = TRUE)
```

>>>>

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
```

@@

 * `group on phylogeny`: the projection of each group's VCV (major axis) on the phylogenetic's VCV (major axis)
 * `species on phylogeny`: the projection of each element within a group (species) onto the phylogenetic's VCV (major axis)
 * `species on group`: the projection of each element within a group (species) onto its own VCV (major axis)

<!-- TODO: add the pictures of the different tests here -->

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE, results = 'hide', message = FALSE, warning = FALSE}
## Run the analyses between groups
super_results_between <- dispRity.covar.projections(superorder_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses between groups
order_results_between <- dispRity.covar.projections(order_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses of groups on the phylogeny
super_results_phylo <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Save all
shapespace_allbirds_lvl_superorder_order_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(shapespace_allbirds_lvl_superorder_order_results_list, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
```

@@

### Projections of groups on the phylogeny:

```{r "starting with boxplots", echo = FALSE, fig.height = 8, fig.width = 8}
## Plot the results
wrap.plot <- function(results, col, main, names, centre.position = TRUE, absolute = FALSE) {

@@

    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    op <- par(mfrow = c(1, n+1))
    ## Plot the first panel
    par(bty = "n", mar = c(5, 14, 4, 0) + 0.1)
    empty <- matrix(0, ncol = length(results[[1]]$disparity))
    colnames(empty) <- names(results[[1]]$disparity)
    boxplot(empty, horizontal = TRUE, las = 2, col = "white", ylim = c(0,0), xaxt = "n", border = "white")
    for(i in 1:n) {
        ## Is position?
        is_position <- names(results)[[i]] == "position"
        par(mar = c(5, 0, 4, 0) + 0.1, bty = "o")
        one.plot(results[[i]], col = col, centre = is_position && centre.position, main = ifelse(i == centre, main, ""), xlab = names_list[i], zero = is_position, absolute = absolute)
    }
    par(op)
}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

```{r, fig.height = 16, fig.width = 8, echo = FALSE}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot the results
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on phylogeny] superorders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on phylogeny] superorders species\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

## Plot the results
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on phylogeny] orders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r "hehehe", echo = FALSE, fig.height = 16, fig.width = 8, eval = FALSE}
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on phylogeny] orders\nspecies onto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
````
</details>

@@

## Plot the results
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on group] superorders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r "beforebefore last", echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on group] superorders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

```{r "before last", fig.height = 16, fig.width = 8, echo = FALSE}

## Plot the results
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on group] oreders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

```{r "done with boxplots", echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on group] oreders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

### Extreme species per super order (species on phylogeny)

@@

### Extreme species per order (species on phylogeny)

@@

### Extreme species per super order (species on group)

@@

### Extreme species per order (species on group)

@@

```{r, echo = FALSE}
## Calculating all the correlations
get.correlations <- function(res, cent.tend = median, exclude.phylo = FALSE, absolute = TRUE) {

@@

    ## Prepare the results
    prep.cor.results <- function(pos, dis, name) {
        ## Run the correlation test
        cor_test <- cor.test(pos, dis)
        ## Prepare the plot title
        plot_main <- paste0(name, ": ", unname(round(cor_test$estimate, 3)), " (p=", unname(round(cor_test$p.value, 3)), ")")
        ## Return the plotting parameters
        return(list(x = pos, y = dis, main = plot_main))
    }

    return(mapply(prep.cor.results, centrals_pos, centrals_dis, as.list(names(centrals_pos)), SIMPLIFY = FALSE))
}

## Plotting one correlation 
plot.correlations <- function(cor.results, col.sub, col.fun = rainbow, legend.pos = "topleft", pt.cex = 0.5, legend.cex = 0.5, ...) {
    dots <- list(...)

    if(missing(col.sub)) {
        col.sub <- NULL
    }

    ## Subbing the colours
    if(!is.null(col.sub)) {
        ## Reset the levels
        col_sub <- as.factor(as.character(col.sub))
        ## Make the empty level grey
        if(any(levels(col_sub) == "")) {
            col_avail <- c("grey", col.fun(length(levels(col_sub))-1))
        } else {
            col_avail <- col.fun(length(levels(col_sub)))
        }
        ## Create the colour vector
        col_vector <- col_avail[as.numeric(col_sub)]

    } else {
        if(is.null(dots$col)) {
            col_vector <- "black"
        } else {
            col_vector <- dots$col
        }
    }

    ## Plotting the results
    plot(cor.results$x, cor.results$y, pch = 19, xlab = "elaboration", ylab = "innovation", main = cor.results$main, cex = pt.cex, col = col_vector)

    if(!is.null(col.sub)) {
        ## Adding the legend
        legend(x = legend.pos, legend = levels(col_sub), pch = 19, cex = legend.cex, col = col_avail)
    }
    return(invisible())
}
```

### Correlations per super order (species on phylogeny)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = TRUE}
## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))
## Calculate the correlations
super_phylo_cor <- get.correlations(results$super_results_phylo)
## Set the vector of legend positions
legends_pos <- c("topleft", "topright", "topright", "topright", "topleft", "topright", "topleft", "topright", "topright")
for(i in 1:length(super_phylo_cor)) {
    plot.correlations(super_phylo_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i])
}
```

### Correlations within super orders (species on group)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = TRUE}
## Getting the subsets for the colouring
col_subs <- split(shapespace$level2, f = shapespace$level1)[-1]
## Calculate the correlations
super_within_cor <- get.correlations(results$super_results_within, exclude.phylo = TRUE)
## Set the vector of legend positions
legends_pos <- c("topleft", "topright", "topright", "topright", "topleft", "topright", "topleft", "topright", "topright")
for(i in 1:length(super_within_cor)) {
    plot.correlations(super_within_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i])
}
```

### Correlations per order (species on phylogeny)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = TRUE}
## Loading the sub-order levels
load(file = "../Data/Processed/shapespace_superorder_lvl_order_suborder.rda")
## Select the data
shapespaces <- lapply(shapespace_superorder_lvl_order_suborder, function(X) X$space)
## Create a level 3
order_space <- do.call(rbind, shapespaces)

## Calculate the correlations
order_phylo_cor <- get.correlations(results$order_results_phylo, exclude.phylo = TRUE)

## Getting the subsets for the colouring
split_tmp <- split(order_space$level2, f = order_space$level1)[-1] ## "Charadriiformes" "Galliformes" not included?
col_subs <- replicate(length(order_phylo_cor), NULL)
col_subs[-c(6,12)] <- split_tmp[match(sort(names(split_tmp)), names(split_tmp))]

## Set the vector of legend positions
for(i in 1:length(order_phylo_cor)) {
    plot.correlations(order_phylo_cor[[i]], col.sub = col_subs[[i]])
}
```

### Correlations within orders (species on group)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = TRUE}
## Calculate the correlations
order_within_cor <- get.correlations(results$order_results_within, exclude.phylo = TRUE)

## Getting the subsets for the colouring
split_tmp <- split(order_space$level2, f = order_space$level1)[-1] ## "Charadriiformes" "Galliformes" not included?
col_subs <- replicate(length(order_within_cor), NULL)
col_subs[-c(6,12)] <- split_tmp[match(sort(names(split_tmp)), names(split_tmp))]

## Set the vector of legend positions
for(i in 1:length(order_within_cor)) {
    plot.correlations(order_within_cor[[i]], col.sub = col_subs[[i]])
}
```
